---
title: "Arima Forecasting"
author: "Sebastian Sternberg"
date: "22 5 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Packages:

```{r}

rm(list = ls())

require(ggplot2)
require(ggthemes)

#set theme for ggplot

theme_set(theme_tufte())

require(magrittr)
require(dplyr)


```



```{r}
source("~/Dropbox/Kaggle/demand_forecasting/Code/helpful_functions.R")

```


# Load the data


```{r}

train_df <- read.csv("~/Dropbox/Kaggle/demand_forecasting/Data/train.csv")
test_df <- read.csv("~/Dropbox/Kaggle/demand_forecasting/Data/test.csv")

head(train_df)
head(test_df)

```


# Explore the data

item is the item id. sales - Number of items sold at a particular store on a particular date. store - Store ID.


```{r}

str(train_df)

train_df$store <- as.factor(train_df$store )
train_df$item <- as.factor(train_df$item )

#do the same for test_df

test_df$store <- as.factor(test_df$store )
test_df$item <- as.factor(test_df$item )

```


## Arima forecasting

ARIMA stands for auto-regressive integrated moving average and is specified by these three order parameters: ($p$, $d$, $d$). 

```{r}

library('forecast')
library('tseries')


train_df$date <- as.Date(train_df$date)

##aggregate sales by date

train_df_plot <- aggregate(sales ~ date, train_df, mean)


ggplot(train_df_plot, aes(date, sales)) + geom_line() + scale_x_date('month')  + ylab(" ") +
            xlab("")


#clean the time serioes

count_ts <- ts(train_df_plot[, c('sales')])

train_df_plot$clean_sales = tsclean(count_ts)

ggplot() +
  geom_line(data = train_df_plot, aes(x = date, y = clean_sales)) + ylab('Cleaned Bicycle Count')


```


```{r}
train_df_plot$sales_ma <- ma(train_df_plot$clean_sales, order=7) # using the clean count with no outliers
train_df_plot$sales_ma30 <- ma(train_df_plot$clean_sales, order=30)


ggplot() +
  geom_line(data = train_df_plot, aes(x = date, y = clean_sales, colour = "Counts")) +
  geom_line(data = train_df_plot, aes(x = date, y = sales_ma,   colour = "Weekly Moving Average"))  +
  geom_line(data = train_df_plot, aes(x = date, y = sales_ma30, colour = "Monthly Moving Average"))  +
  ylab('Bicycle Count')

```


## Decompose the data into blocks

calculate seasonal component of the data using stl(). STL is a flexible function for decomposing and forecasting the series. It calculates the seasonal component of the series using smoothing, and adjusts the original series by subtracting seasonality in two simple lines:


```{r}

count_ma <- ts(na.omit(train_df_plot$sales_ma), frequency=30)
decomp <- stl(count_ma, s.window="periodic")
deseasonal_cnt <- seasadj(decomp)
plot(decomp)


#test for stationarity:

adf.test(count_ma, alternative = "stationary")

```


## Autocorrelations and Choosing Model Order


```{r}
Acf(count_ma, main='')

Pacf(count_ma, main='')


count_d1 = diff(deseasonal_cnt, differences = 1)
plot(count_d1)
adf.test(count_d1, alternative = "stationary")

```


```{r}

Acf(count_d1, main='ACF for Differenced Series')
Pacf(count_d1, main='PACF for Differenced Series')

```


## Fit the model:


```{r}

auto.arima(deseasonal_cnt, seasonal=FALSE)




```

## Evaluate and iterate

```{r}

fit<-auto.arima(deseasonal_cnt, seasonal=FALSE)
tsdisplay(residuals(fit), lag.max=45, main='(4,1,5) Model Residuals')

```


```{r}

fit2 <- arima(deseasonal_cnt, order=c(4,1,7))
tsdisplay(residuals(fit), lag.max=15)


fcast <- forecast(fit2, h=30)
plot(fcast)


###using holdout set:

hold <- window(ts(deseasonal_cnt), start=1454)

fit_no_holdout = auto.arima(ts(deseasonal_cnt[-c((length(deseasonal_cnt)- 30) :length(deseasonal_cnt))]))

fcast_no_holdout <- forecast(fit_no_holdout,h=30)
plot(fcast_no_holdout, main=" ")
lines(ts(deseasonal_cnt))


```



## Using seasonality:


```{r}

fit_w_seasonality = auto.arima(deseasonal_cnt, seasonal=TRUE)
fit_w_seasonality

seas_fcast <- forecast(fit_w_seasonality, h=30)
plot(seas_fcast)

tsdisplay(residuals(seas_fcast), lag.max=45)


fcarima <- auto.arima(train_df_plot[, c("sales")])
fcarima %>% forecast(h = 90) %>% autoplot()
checkresiduals(fcarima)

seas_fcast <- forecast(fcarima, h=90)

```

## Now make an individual model for all the different stores

```{r}

train_df_reshaped <- train_df %>% reshape(.,  idvar =c('store', 'item'), direction = 'wide', timevar = 'date')

head(train_df_reshaped)
tail(train_df_reshaped)

```

For each store, we have now for each item them sales per day. 

Write a function so that we can apply autoarima for each of the stores and items separately. Thus, we run 500 time series models. 


```{r}
        m1=  auto.arima((log1p(train_df_reshaped[1, 3:1828])),max.p = 15, max.q = 15, max.P = 2, max.Q = 2, max.order = 5, max.d = 1, max.D = 1, seasonal=TRUE)

bla <- (log1p(train_df_reshaped[1, 3:10]))

 auto.arima(log1p(train_df_reshaped[1, 3:ncol(train_df_reshaped)]))
#log1p(train_df_reshaped[1, 3:ncol(train_df_reshaped)])
bla <- log1p(train_df_reshaped[1, 3:ncol(train_df_reshaped)]) %>% as.vector()

sum(is.numeric(bla))

dim(train_df_reshaped)
some_function <- function(x){
        m1 <-  auto.arima((log1p(x[3:ncol(train_df_reshaped)])), 
                          max.p = 15, 
                          max.q = 15, 
                          max.P = 2, 
                          max.Q = 2, 
                          max.order = 5, 
                          max.d = 1, 
                          max.D = 1, 
                          seasonal=TRUE)
        f1 <- forecast(m1,h=90)
        tmp <- as.data.frame(f1)
        names(tmp)=c("Forecast","Lo_80","Hi_80","Lo_95","Hi_95")
        tmp['store'] = x['store']
        tmp['item']= x['item']
        head(tmp)
        return(tmp)
}
res1= apply(train_df_reshaped,1,some_function)

print("--- done ")
res2 = reduce(res1,bind_rows)
subm = read.csv("../input/sample_submission.csv")
res3 = cbind(subm,res2)
names(res3) = c("id","Forecast" ,"sales", "Lo_80",    "Hi_80",    "Lo_95"    ,"Hi_95" ,"store"   , "item")
res3$sales = expm1(res3$sales)
res5 = res3 %>% select(id,sales)
write.csv(res5,"auto_arima_log.csv",row.names=FALSE)

```























