---
title: "EDA Demand Forecasting Kaggle"
author: "Sebastian Sternberg"
date: "21 5 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

sessionInfo()

```

Packages:

```{r}

rm(list = ls())

require(ggplot2)
require(ggthemes)

#set theme for ggplot

theme_set(theme_tufte())

require(magrittr)
require(dplyr)


```



```{r}
source("~/Dropbox/Kaggle/demand_forecasting/Code/helpful_functions.R")

```


# Load the data


```{r}

train_df <- read.csv("~/Dropbox/Kaggle/demand_forecasting/Data/train.csv")
test_df <- read.csv("~/Dropbox/Kaggle/demand_forecasting/Data/test.csv")

head(train_df)
head(test_df)

```


# Explore the data

item is the item id. sales - Number of items sold at a particular store on a particular date. store - Store ID.


```{r}

str(train_df)

train_df$store <- as.factor(train_df$store )
train_df$item <- as.factor(train_df$item )

#do the same for test_df

test_df$store <- as.factor(test_df$store )
test_df$item <- as.factor(test_df$item )

```



```{r}

ggplot(data = train_df) +
  geom_bar(mapping = aes(x = store))

```

each store exactly the same amount of observations. 


```{r}

ggplot(data = train_df) +
  geom_bar(mapping = aes(x = item))





```

That fits the data description: sales for 50 different items at 10 different stores.


```{r}

ggplot(data = train_df, aes(x=sales)) + 
  geom_density() + theme_tufte()

```

See how the sales vary over the different stores:

```{r}

ggplot(data = train_df, mapping = aes(x = store, y = sales)) +
  geom_boxplot() + theme_tufte()

#reordered version:

library(forcats)
ggplot(train_df, aes(x = fct_reorder(store, sales, fun = median, .desc =TRUE), y = sales)) + 
        geom_boxplot() + xlab("Store") 


#get the group means
train_df %>%
  group_by(store) %>%
  summarise(sales = mean(sales)) %>%
  arrange(sales)

#lookt at density over stores:
ggplot(train_df) + geom_density(aes(x = sales)) + facet_wrap(~store) + theme_tufte()

```

Store 2 and 8 have the most sales (67 and 64) overall, and store 7 and 6 the least sales.



See how the sales vary over the different items:

```{r}

ggplot(train_df, aes(x = fct_reorder(item, sales, fun = median, .desc =TRUE), y = sales)) + 
        geom_boxplot() + xlab("Items")

#get the group means
train_df %>%
  group_by(item) %>%
  summarise(sales = mean(sales)) %>%
  arrange(item)

#lookt at density over stores:
ggplot(train_df) + geom_density(aes(x = sales)) + facet_wrap(~item) + theme_tufte()
```

Item 15, 28, 13, 18 are the best selling items overall


Check Nas:

```{r}

plot_missing(train_df)

```


# Exploring the time dimension:

```{r}
require(zoo)
require(lubridate)

# Extraction of Year and Month of Year :
train_df$Year <- year(train_df$date)        #returns the year from date i.e. 2013, 2014 etc.
train_df$MonthYear <- as.yearmon(train_df$date)

#aggregate the sales per month year

train_df_monthyear <- train_df %>% 
  group_by(MonthYear) %>%
  summarise(sales_monthyear_mean = mean(sales))


head(train_df_monthyear)

#plot time series by month and year
library(scales)

ggplot(train_df_monthyear, aes(x = as.POSIXct(MonthYear), y = sales_monthyear_mean)) + 
  geom_line() + 
  stat_smooth(
  method = "loess"
  ) + 
  scale_x_datetime(labels = date_format("%Y-%m"), 
                   breaks = date_breaks("months")) + 
  theme(axis.text.x = element_text(angle = 45))


```


```{r}

train_df_monthyear$rate <- c(0, 100*diff(train_df_monthyear$sales_monthyear_mean)/train_df_monthyear[-nrow(train_df_monthyear),]$sales_monthyear_mean)

```



# A very simply OLS baseline


```{r}

summary(m1_lm <- lm(sales ~ store + item, data = train_df))

predict_lm <- predict(m1_lm, test_df)

df_prediction <- cbind.data.frame(id = test_df$id, sales = predict_lm)

write.csv(df_prediction, file = "sample_submission.csv", row.names = F)

```

The simple OLS achieves a score of 48.78430 on the submission file.


# Start Forecasting with Prophet (from Facebook)



prepare the data: The ds column should be YYYY-MM-DD for a date, or YYYY-MM-DD HH:MM:SS for a timestamp.

```{r}
library(prophet)
require(tibble)

train_df_prophet <- train_df %>% filter(Year < 2014) %>% 
  select(ds = date, store, item, y = sales)
#rename date to ds. the format is already the necessary one

#subset for testing only one year:

m <- prophet(train_df_prophet)

future <- make_future_dataframe(m, periods = 365)
forecast <- predict(m, future)

plot(m, forecast)
```







